resistanceModelNotesAndy

### 2015-04-29 resistanceModelNotesAndy.Rmd
progress notes

**27/4/15** 9.15-13.15 4hrs

@@

@@

1. continue with getting classification trees working in sensiAn1

1. update R, new lengths() function may be useful for me
but doesn't find latest version yet
library(installr)
updateR()

@@

1. from the first stage of the sensitivity analysis I'll want to output a dataframe with 1 row per replicate containing both the inputs and the outputs. Getting it to put the inputs into a dataframe like this would be a good start. I think this may just be a case of transposing the input object ?

1. try out classification trees code from Susana

@@

1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
1. edit runModel carefully
1. check out the part about running sequential insecticide scenarios

1. spend some time developing my understanding of how the model works
1. submit tilde bug report
1. re-look at R popgen packages when I've got more of a handle on what we are doing.

### suggestion from Beth
a solution may be to provide a .csv file with the Curtis scenarios set up
and then the need to set the calibration number and params.csv in the R script can be removed altogether

@@

```
colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
                        "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
                        "RR1SS2", "RR1RS2", "RR1RR2")
```

>>>>

**27/4/15** 9.15-13.15 4hrs 1.45-2.45 5hrs

@@

**28/4/15** 10-14.30 4.5hrs 3-6 7.5hrs

reading Ians recent manuscript update.
I think this from p13 is key on what I need to do with the analysis :

** The simulation output was the ratio of  the time (in generation) for resistance  to the ‘new’ insecticide  to reach a given allele threshold frequency when the ‘new’ insecticide was deployed as combination, compared to the time taken to reach the same threshold if deployed on its own; a ratio >1 indicates that combination is better. Five threshold frequencies were investigated: 0.05, 0.1, 0.2, 0.5, 0.8. **

In order to do that I'd need to repeat each scenario for a single & combined insecticide treatment. Single is just the new rarer insecticide, Combined also has an older insecticide to which initial resistance is more common.

Searching for the Curtis paper. Cited by 181 papers according to Google.
https://www.google.co.uk/webhp?sourceid=chrome-instant&ion=1&espv=2&ie=UTF-8#q=%22theoretical%20models%20of%20the%20use%20of%20insecticide%20mixtures%22

https://books.google.co.uk/books?id=ooHjBwAAQBAJ&pg=PA177&lpg=PA177&dq=%22theoretical+models+of+the+use+of+insecticide+mixtures%22&source=bl&ots=b3CPTQc5LL&sig=FCB5MCvX1xUjtfgog1z_dq1pH1Y&hl=en&sa=X&ei=bXw_VbriFoWpsAGnvYG4Ag&ved=0CEEQ6AEwBA#v=onepage&q=Curtis&f=false

Pesticide Resistance in Arthropods (1990) edited by Richard Roush, Bruce E. Tabashnik

these 2 chapters look useful :
Roush, R. T., and J. C. Daly. 1990. The role of population genetics in resistance research
and management, pp. 97-152.In: Pesticide resistance in arthropods. R. T. Roush and
B. E. tabashnik (Eds.) Chapman and Hall, New York.

Tabashnik, B. E. 1990. Modeling and evaluation of resistance management tactics, pp.
153-182. In: Pesticide resistance in arthropods. R. T. Roush and B. E. Tabashnik
(Eds.) Chapman and Hall, New York.

page 131 talks about insecticide mixtures. Refers to some modelling by Roush (1989a)

Roush, R. T. 1989. Designing resistance management programs: How can you choose?
Pestic. Sci. 26:423-441.

p467 in Insect Resistance Management: Biology, Economics, and Prediction edited by David W. Onstad (2013)
seems to have some simulations addressing similar issues to us.

https://books.google.co.uk/books?id=6hp384ZH0_kC&pg=PA453&dq=%22Modeling+and+evaluation+of+resistance+management+tactics%22&source=gbs_toc_r&cad=4#v=onepage&q=%22Modeling%20and%20evaluation%20of%20resistance%20management%20tactics%22&f=false

Now have sorted passing user args to rpart in sensiAn1(). It does produce reasonable looking trees just by e.g.

`inputAndResults <- sensiAn1(100, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))`

**29/4/15** 10.30-12.30 2hrs 3.15-

working on refactoring variables to arrays

@@

1. is there a way I can easily run single & combined deployments for the same scenario ?
* is it `niche_00` and `niche_AB`
* or in `a.m_AB` insecticide exposure male hi1 hi2
* do I also need to change "Starting frequency of resistance" ?
* perhaps I can set up a simple example input to send to Beth ?
* or ask Beth to send me the input.csv used to produce tables 8 & 9 in 'Brief_full-genotype-tables.pdf'
* maybe comparing the files in full-dom_CD.csv & SD in C:\Dropbox\Levick et al IR manuscript will help me.

1. before May 15th move Spark and Glimmer shiny apps

1. go through the paper and check which bits relate to which in the code. Maybe create a table ? Could I create the table from R ? but actually much of the functionality is still in runModel() and needs to be put into functions.

1. Maybe create a model diagram perhaps from diagrammer ?

1. maybe separate out classification trees from sensiAn1

@@

@@

1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
1. edit/refactor runModel carefully
1. check out the part about running sequential insecticide scenarios

1. spend some time developing my understanding of how the model works
1. submit tilde bug report

### suggestion from Beth
could provide a .csv file with the Curtis scenarios set up
and then the need to set the calibration number and params.csv in the R script can be removed altogether

### How best to refactor variables to arrays ?
Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.
cis/trans makes it slightly tricky, but not all variables have cis/trans
only used for genotype frequencies (f) and in output matrices fitness & genotype 
in the wild in Beths code :
```
fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
```

e.g. W doesn't have cis/trans
I could refactor
```{r}
        # relaxed selection fitnesses
        ## Males
        W.m.SS1SS2 <- 0.1 
        W.m.SS1RS2 <- 0.1
        W.m.SS1RR2 <- 0.1

        W.m.RS1SS2 <- 0.1
        W.m.RS1RS2 <- 0.1  
        W.m.RS1RR2 <- 0.1

        W.m.RR1SS2 <- 0.1 
        W.m.RR1RS2 <- 0.1 
        W.m.RR1RR2 <- 0.1

        ## Female
        W.f.SS1SS2 <- 0.1
        W.f.SS1RS2 <- 0.1
        W.f.SS1RR2 <- 0.1

        W.f.RS1SS2 <- 0.1
        W.f.RS1RS2 <- 0.1 
        W.f.RS1RR2 <- 0.1

        W.f.RR1SS2 <- 0.1
        W.f.RR1RS2 <- 0.1 
        W.f.RR1RR2 <- 0.1
# to :
#to create an array a[sex][locus1][locus2]
namesLoci <- c('SS','RS','RR')
locus1 <- paste0(namesLoci,'1')
locus2 <- paste0(namesLoci,'2')
sex <- c("f","m")

dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
dim1 <- sapply(dimnames1, function(x) length(x))
W <- array(0,dim=dim1, dimnames=dimnames1)

#to fill al elements as above
W <- 0.1
#check length
length(W)
#to access one element
W['m','SS1','RR2']
#to check that males total 1
sum(a['m',,])
```

@@

```
colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
                        "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
                        "RR1SS2", "RR1RS2", "RR1RR2")
```
#### inputs
** a ** exposure to insecticide
** phi ** fitness of one locus (baseline)
** W.SS1_00 & W.SS2_00** fitness of SS in no insecticide
** h ** dominance coefficient
** s ** selection coefficient
** z ** fitness cost of resistance allele in no insecticide
** niche_ ** insecticide niche toggle

#### calculated
** W ** fitnesses by niche and sex
** f ** genotype frequency
** fs ** genotype frequency following selection
** G ** gametes

Different W fitness variables
`W.RR1_00 : [locus][niche]` single locus
`W.RR1SS2_0b :   [locus1][locus2][niche]` niche
`W.m.SS1SS2 : [sex][locus1][locus2]` sex

Question.
What if `a.m_A0 > 0` but `niche_A0 == 0` ?
The ifelse clauses allow niches to be toggled on/off, i.e. a fitness can be given for A0
but if only the niche A,B is toggled on, the fitness scores for A0 and Ab will be set to 0
